---
title: "data_cleaning"
format: html
edtiro: visual
---

# Libraries

```{r}
library("tidyverse")
```

IF "YOUNG" IN DATA - IGNORE/ELIMINATE

# Cleaning table 1

Discard iBAQ.

LFQ intensity is the relative abundance of proteins with one another. Also keep the information on the right side of the table (peptides, etc.).

Clean NA's in protein names.

Figure out how to handle unique names on both gene and protein names.

Eliminate PEP column and Mol. weight.

Rename columns to have standard names

```{r}
#| echo: false
dirty_1 <- read_tsv(file = "../data/dirty_data/supp_1.tsv")
```

# Cleaning table 2

Eliminate standard deviation part.

Keep things on the right too (peptides, etc).

Keep only logLFQ.

Separate data into different tables (isoalted cells, brain regions and cerebellum development)

Eliminate PEP column and Mol. weight.

Rename columns to have standard names

```{r}
#| echo: false
dirty_2 <- read_tsv(file = "../data/dirty_data/supp_2.tsv")
```

# Cleaning table 3
Before removing NaN values, the data has: 19,501 entries
After removing NaN values, the data has: 11,062 entries

Rename columns so they have more standard names

```{r}
#| echo: false
library("dplyr")
library("janitor")

dirty_3 <- read_tsv(file = "../data/dirty_data/supp_3.tsv")

#Transforming NaNs into NAs to drop them and cleaning column names
clean_3 <- drop_na(
  mutate(
    clean_names(dirty_3),
    across(everything(), 
    ~replace(., is.nan(.), NA))
  )
)  
```

# Cleaning table 5

Eliminate PEP column and Mol. weight.

Rename columns to have standard names.

Remove Protein IDs and Majority Protein IDs

```{r}
#| echo: false
dirty_5 <- read_tsv(file = "../data/dirty_data/supp_5.tsv")
```
